---
title: "p8105 Homework 2"
author: Tory Lynch 
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(p8105.datasets)
library(ggplot2)
```

#### Problem 1

##### Import CSV file and clean data 
```{r}
MTA_data <- read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv", 
                     col_types = "cccddccccccciiiicccccclclcccddcc") %>%
            janitor::clean_names() %>%
            dplyr::select(line:entry, vending, ada) %>%
            mutate(entry = as.logical(ifelse(entry == "YES", TRUE, FALSE))) 

skimr::skim(MTA_data)
```

```{r, include = FALSE}
dimensions <- dim(MTA_data)
row <- dimensions[1]
column <- dimensions[2]
```

##### Description of dataset 
The dataset contains information about all of the NYC MTA subway station lines, including the station stops, locations (latitude, longitude, cross streets) and routes served (i.e. the numbers and/or letters of the trains on each line). Many of the variables address the accessiblity of the station line; these data describe the entrance and exit type for each station, whether or not they are staffed, have metro card vending machines, and are ADA-compliant. The dataset also includes information about the station entrances (lat, long, corner position) that is separate from the station location. The data cleaning process involved restricting the dataset from 32 to 19 variables. I removed the variables that described the entrance location (lat, long, corner position), whether the stations were exit only, and if the stations were staffed. I also converted the entry variable, which identifies whether a station has an entrance, from a character to a logical variable. The dimensions of the data are `r row` by `r column`. The data are tidier because I arranged them in a clear format by selecting the relevant variables and by converting a variable into an easier-to-use format. 

##### Questions 
```{r}
#Question 1
table_distinct_stations <- distinct(MTA_data) 
num_distinct_stations <- as.integer(count(table_distinct_stations))

#Question 2 
proportion_ada_compliant <- (round(sum(table_distinct_stations$ada == "TRUE") / num_distinct_stations, 3))*100

#Question 3 
prop_entry_no_vending <- (round((sum(table_distinct_stations$vending == "NO" & 
          table_distinct_stations$entry == "TRUE") /sum(table_distinct_stations$vending == "NO")), 3))*100
```

There are `r num_distinct_stations` distinct stations, of which `r proportion_ada_compliant`% are ADA-compliant. Among the stations without vending, `r prop_entry_no_vending`% allow entrance. 

##### Reformat data
```{r}
stations_name_and_route <- gather(table_distinct_stations, key = route, value = line, route1: route11) %>%
  filter(!is.na(line)) %>%
  select(-route)

stations_per_line <- stations_name_and_route %>%
  group_by(line, ada) %>%
  tally()
total_a_train <- 34+57
percent_ada_compliant <- (34 / total_a_train)*100
```

There are `r total_a_train` distinct stations that serve the A line, `r percent_ada_compliant`% of which are ADA compliant.

#### Problem 2

##### Import Excel file and clean data
```{r}
#Trashwheel data
trashwheel_data <- read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                              sheet = "Mr. Trash Wheel", 
                              range = cell_cols("A:N")) %>% 
                   janitor::clean_names() %>% 
                   filter(!is.na(dumpster)) %>% 
                   mutate(sports_balls = as.integer(round(sports_balls, 0)))

#2017 precipitation data  
precip_2017 <- read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                              sheet = "2017 Precipitation", 
                              skip = 2, 
                              col_names = c("Month", "Precip")) %>%  
               filter(!is.na(Month)) %>% 
               mutate(Year = 2017)

#2016 precipitation data 
precip_2016 <- read_excel("data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
                              sheet = "2016 Precipitation", 
                              skip = 2, 
                              col_names = c("Month", "Precip")) %>%  
               filter(!is.na(Month)) %>% 
               mutate(Year = 2016)                            

#Combine precipitation data 
total_precip_data <- bind_rows(precip_2016, precip_2017) %>% 
               mutate(month = as.character(month.name[Month])) %>% 
               select(-Month)
```

```{r}
skimr::skim(trashwheel_data)
skimr::skim(total_precip_data)

total_2017_precip <- total_precip_data %>% 
  group_by(Year) %>% 
  tally(Precip)

total_precip_data <- total_precip_data %>% 
  mutate()

sportsballs_by_year <- trashwheel_data %>% 
  group_by(year) %>% 
  summarise(n = median(sports_balls))

num_trashwheels <- n_distinct(trashwheel_data$dumpster)
num_precip_months <- n_distinct(total_precip_data)
```

##### Description of dataset 
These datasets contain information about the Healthy Harbor Initiative in Baltimore, specifically the amount and type of trash collected by the trashwheel, and monthly precipitation in the surrounding area. The trashwheel dataset contains `r num_trashwheels` obsevations, each of which represents a dumpster-load of trash collected by the wheel. The dataset provides a breakdown of the type (e.g. plastic bottles, cigarette butts) and number of pieces of trash trapped by the trashwheel per dumpster-load each month between 2014 and 2018. The dataset also contains a variable that records how many homes were powered by the incineration of the trash; this variable is important because it underscores the mission behind the program, which is clean up the harbor while providing a "use" for the trash. The precipitation dataset has `r num_precip_months` observations, or months of data; these data are important because the amount of trash collected depends on the amount of precipitation and the associated runoff. The total precipitation in 2017 was 32.93 inches and the median number of sports balls per dumpster in 2016 was eight. 

#### Problem 3 

##### Import data
```{r}
 data("brfss_smart2010")
 head(brfss_smart2010)
 
 brfss_data <- filter(brfss_smart2010, Topic == "Overall Health") %>% 
   janitor::clean_names() %>% 
   dplyr::select(-(class:question), -(sample_size), -(confidence_limit_low:geo_location)) %>% 
   spread(key = response, data_value) %>% 
   janitor::clean_names() %>% 
   mutate(proportion_excellent_or_good = (excellent + very_good)/(excellent+very_good+good+fair+poor)) %>% 
   mutate(proportion_excellent = (excellent/(excellent+very_good+good+fair+poor)))
```

##### Questions 
```{r}
#Number of unique locations
unique_locations <- n_distinct(brfss_data$locationdesc)

#Number of locations per state 
num_states <- brfss_data %>% 
  group_by(locationabbr) %>% 
  count()

#Median of excellent response variable in 2002
median_excellentRV_2002 <- brfss_data %>% 
  drop_na() %>% 
  group_by(year) %>% 
  summarise(n = median(excellent))
```

There are `r unique_locations` unique locations included in the dataset. Each state, plus Washington, DC, is represented and New Jersey is the most observed state. The median of the excellent response variable in 2002 is 23.6, which means that the median percentage reporting excellent health is 23.6%.  

##### Plots
```{r}
#Histogram
ggplot(brfss_data, aes(x = excellent)) + geom_histogram(data=subset(brfss_data, year == 2002), na.rm = TRUE, bins = 30,
      fill = "blue", alpha = .4) + labs(title = "Histogram of Proportion of People Reporting Excellent Health in 2002") + 
      theme(plot.title = element_text(hjust = 0.5))

#Scatterplot
brfss_data %>% 
  mutate(counties = ifelse(locationdesc == "NY - New York County", 1, ifelse(locationdesc == "NY - Queens County", 2,0))) %>% 
  filter(counties > 0) %>% 
  ggplot(aes(x = year, y = proportion_excellent)) + geom_point(aes(color = factor(counties)))  + 
      theme(legend.title = element_text(colour = "black", size = 10)) + labs(title = "Proportion Reporting Excellent Health in New York 
      County and Queens County between 2002 - 2010", x = "Year", y = "Proportion in excellent health") + 
      scale_color_manual(labels = c("New York County", "Queens County"), values = c('red', 'blue'), name = "NY Counties") + 
      theme(plot.title = element_text(hjust = 0.5))
```

